A Symbolic Decision Procedure for GKAT

### 2024-05-21 notes.tex
minor update that doesn't matter that much

Bisimilarity implies language sematnical equivalence \[s ∼ t ⟹ ⟦s⟧ = ⟦t⟧\], however the converse doesn't hold unless we are working in a \emph{normal} GKAT coalgebra.
It is well known that two states in a normal GKAT coalgebra if and only if they have the same language semantics \[s ∼ t ⟺ ⟦s⟧ = ⟦t⟧.\]
However, instead of normalizing automatons, we can in fact, check for normality on the fly, this is important to design a derivatives based decision algorithm.

@@

Finally, to construct a normalized bisimulation using programs, we will need the following inductive construction theorem
\begin{theorem}[Inductive Construction]
    Given two DKCT \(𝒮\) and \(𝒯\), and two of their elements \(s ∈ 𝒮\) and \(t ∈ 𝒯\),
    there exists a normalized bisimulation \(s ∼ t\) between \(⟨s⟩\) and \(⟨t⟩\), if and only if either both \(s\) and \(t\) are dead or for all \(α ∈ \At\), all of the following holds:
    \begin{enumerate}
        \item \(δ_{𝒮}(s, α) = \accept ⟺ δ_{𝒯}(t, α) = \accept\);
        \item\label{itm: transition condition} If \(δ_{𝒮}(s, α) = (s', p)\) and \(δ_{𝒯}(t, α) = (t', p)\), then and there exists a normalized bisimulation \({∼_{s',t'}}\) on \(⟨s'⟩\) and \(⟨t'⟩\), s.t. \(s' ∼_{s',t'} t'\);
        \item If \(δ_{𝒮}(s, α) = (s', p)\) and \(δ_{𝒯}(t, α) = (t', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead;
        \item \(s\) reject \(α\) or transition to a dead state via \(α\) if and only if \(t\) rejects \(α\) or transition to a dead state via \(α\).
    \end{enumerate}
\end{theorem}

@@

@@

>>>>

Bisimilarity implies language semantical equivalence \[s ∼ t ⟹ ⟦s⟧ = ⟦t⟧\], however the converse doesn't hold unless we are working in a \emph{normal} GKAT coalgebra.
It is well known that two states in a normal GKAT coalgebra if and only if they have the same language semantics \[s ∼ t ⟺ ⟦s⟧ = ⟦t⟧.\]
However, instead of normalizing automatons, we can in fact, check for normality on the fly, this is important to design a derivatives based decision algorithm.

@@

Finally, to construct a normalized bisimulation using programs, we will need the following inductive construction theorem
\begin{theorem}[Inductive Construction]
    Given two DKCT \(𝒮\) and \(𝒯\), and two of their elements \(s ∈ 𝒮\) and \(t ∈ 𝒯\),
    there exists a normalized bisimulation equivalence \({∼} ⊆ ⟨s⟩ × ⟨t⟩\) s.t. \(s ∼ t\), if and only if either both \(s\) and \(t\) are dead or for all \(α ∈ \At\), all of the following holds:
    \begin{enumerate}
        \item \(δ_{𝒮}(s, α) = \accept ⟺ δ_{𝒯}(t, α) = \accept\);
        \item\label{itm: transition condition} If \(δ_{𝒮}(s, α) = (s', p)\) and \(δ_{𝒯}(t, α) = (t', p)\), then and there exists a normalized bisimulation \({∼_{s',t'}}\) on \(⟨s'⟩\) and \(⟨t'⟩\), s.t. \(s' ∼_{s',t'} t'\);
        \item If \(δ_{𝒮}(s, α) = (s', p)\) and \(δ_{𝒯}(t, α) = (t', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead;
        \item \(s\) reject \(α\) or transition to a dead state via \(α\) if and only if \(t\) rejects \(α\) or transition to a dead state via \(α\).
    \end{enumerate}
\end{theorem}

@@

\begin{corollary}
    when constructing \(∼\) notice that 
\end{corollary}

@@

\section{Symbolic Coalgebra}

\begin{theorem}
    Given two states \(s, t\) there exists a normalized bisimulation \({∼} ⊆ ⟨s⟩ × ⟨t⟩\) if and only if the following is true 
    \begin{itemize}
        \item \(ϵ(s) = ϵ(t)\)
        \item for all \((b, s', p) ∈ δ(s)\), \(b ∧ \)
    \end{itemize}
\end{theorem}
